---
title: 'Practical Machine Learning Assignment'
author: 'Author: UltraMagnus'
output:
  html_document:
    keep_md: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
---
# Overview

# Loading the Data  

Reading the Data from csv files 'Tdata' contain the traning data and Sdata contains the 20 test case to be submitted.  

```{r}
dirpath="/home/lotus/Data_Science/08_Practical Machine Learning/GIT/"
Tdata<-read.table(paste0(dirpath,"pml-training.csv"), header = TRUE, sep = ",")
Sdata<-read.table(paste0(dirpath,"pml-testing.csv"), header = TRUE, sep = ",")
```

# Cleaning the Data  

1. Calculating the index of columns which contains the NA values, rmCol contains the 
index of those columns.
```{r}
library(dplyr)
i<-0
rmCol<-NULL
for(i in 1:ncol(Sdata))
{
  if(sum(is.na(Sdata[, i]))  > 0)
		rmCol<-c(rmCol,i)
}
```

removing column containing NA values from both 

```{r}
Tdata<-select(Tdata, -rmCol)
Sdata<-select(Sdata, -rmCol)
```

removing unuseful colums  
```{r}
Tdata<-select(Tdata, -c(X,user_name, cvtd_timestamp, new_window, num_window))
Sdata<-select(Sdata, -c(X,user_name, cvtd_timestamp, new_window, num_window))
dim(Tdata)
dim(Sdata)
```

so finally both contains 55 columns  

# Creating Traning and Test Set
creating the Traning and Test set 70% for Training and 30% for Testing. 
```{r}
library(caret)
set.seed(12321)
testIndex = createDataPartition(Tdata$classe, p = 0.70,list=FALSE)
training = Tdata[-testIndex,]
testing = Tdata[testIndex,]
```

# Training  

Training the training data using randomForest.  

```{r}
library(randomForest)
modFit <- randomForest(classe ~ .,data=training)
```


#Testing 
Testing the model, 55th column is not used as it is 'problem_id' variable  
```{r}
confusionMatrix(testing$classe, predict(modFit,testing[,-55]))
```

we can see accurary is 99.16% which is quite statisfactory.  so we can use above model for final test cases for submission.  

On 20 Test Cases Results are  
```{r}
result<-predict(modFit,Sdata[,-55])
result
```

all test cases passed.
